---
output: github_document
bibliography: readme.bib
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(18052020)
```

# mcmcensemble

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version-ago/mcmcensemble)](https://CRAN.R-project.org/package=mcmcensemble)
[![R build status](https://github.com/Bisaloo/mcmcensemble/workflows/R-CMD-check/badge.svg)](https://github.com/Bisaloo/mcmcensemble/actions)
[![Codecov test coverage](https://codecov.io/gh/Bisaloo/mcmcensemble/branch/master/graph/badge.svg)](https://codecov.io/gh/Bisaloo/mcmcensemble?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)

This R package provides ensemble samplers for affine-invariant Monte Carlo
Markov Chain, which allow a faster convergence for badly scaled estimation
problems. Two samplers are proposed: the 'differential.evolution' sampler from
@terBraak2008 and the 'stretch' sampler from @Goodman2010.

For theoretical background about Ensemble MCMC (what are the benefits over
simple MCMC? How do they work? What are the pitfalls?), please refer for example to 
[this lecture](https://doi.org/10.26207/46za-m573) from Eric B. Ford (Penn State).

## Installation

You can install the stable version of this package from 
[CRAN](https://cran.r-project.org/package=mcmcensemble):

```{r, eval = FALSE}
install.packages("mcmcensemble")
```

or the development version from [GitHub](https://github.com/bisaloo), via my 
[r-universe](https://bisaloo.r-universe.dev/):

```{r, eval = FALSE}
install.packages("mcmcensemble", repos = "https://bisaloo.r-universe.dev")
```

## Usage

```{r, include = FALSE}
set.seed(20210220)
```

```{r}
library(mcmcensemble)

## a log-pdf to sample from
p.log <- function(x) {
  B <- 0.03 # controls 'bananacity'
  -x[1]^2 / 200 - 1/2 * (x[2] + B * x[1]^2 - 100 * B)^2
}


## use stretch move
res1 <- MCMCEnsemble(
  p.log,
  lower.inits = c(a = 0, b = 0), upper.inits = c(a = 1, b = 1),
  max.iter = 3000, n.walkers = 10, method = "stretch"
)

attr(res1, "ensemble.sampler")

str(res1)
```

If the [coda](https://cran.r-project.org/package=coda) package is installed,
you can then use the `coda = TRUE` argument to get objects of class `mcmc.list`.
The coda package then allows you to call `summary()` and `plot()` to get 
informative and nicely formatted results and plots:

```{r example-stretch, dev = 'svglite'}
## use stretch move, return samples as 'coda' object
res2 <- MCMCEnsemble(
  p.log,
  lower.inits = c(a = 0, b = 0), upper.inits = c(a = 1, b = 1),
  max.iter = 3000, n.walkers = 10, method = "stretch", coda = TRUE
)

attr(res2, "ensemble.sampler")

summary(res2$samples)
plot(res2$samples)
```

```{r example-de, dev = 'svglite'}
## use different evolution move, return samples as 'coda' object
res3 <- MCMCEnsemble(
  p.log,
  lower.inits = c(a = 0, b = 0), upper.inits = c(a = 1, b = 1),
  max.iter = 3000, n.walkers = 10, 
  method = "differential.evolution", coda = TRUE
)

attr(res3, "ensemble.sampler")

summary(res3$samples)
plot(res3$samples)
```

## Parallel processing

This package is set up to allow transparent parallel processing when requested 
by the user thanks to the framework provided by the 
[future](https://cran.r-project.org/package=future) package. To enable parallel 
processing, you must run:

```{r, eval = FALSE}
future::plan("multiprocess")
```

at the start of your session.

## Similar projects

The Goodman-Weare 'stretch' sampler is also available in the [tonic R package](https://github.com/svdataman/tonic).
 
The methods used in this package also have (independent) implementations in 
other languages:

- [emcee v3: A Python ensemble sampling toolkit for affine-invariant MCMC](https://doi.org/10.21105/joss.01864)
- [GWMCMC which implements the Goodman-Weare 'stretch' sampler in Matlab](https://github.com/grinsted/gwmcmc)

## Who is talking about this package?

- [R View from October 2020](https://rviews.rstudio.com/2020/11/19/october-2020-top-40-new-cran-packages/)

## References
