---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(18052020)
```

# mcmcensemble

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version-ago/mcmcensemble)](https://CRAN.R-project.org/package=mcmcensemble)
[![R build status](https://github.com/Bisaloo/mcmcensemble/workflows/R-CMD-check/badge.svg)](https://github.com/Bisaloo/mcmcensemble/actions)
[![Codecov test coverage](https://codecov.io/gh/Bisaloo/mcmcensemble/branch/master/graph/badge.svg)](https://codecov.io/gh/Bisaloo/mcmcensemble?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
<!-- badges: end -->

This R package provides ensemble samplers for affine-invariant Monte Carlo
Markov Chain, which allow a faster convergence for badly scaled estimation
problems. Two samplers are proposed: the 'differential.evolution' sampler from
[ter Braak and Vrugt (2008)](https://doi.org/10.1007/s11222-008-9104-9) and the
'stretch' sampler from [Goodman and Weare
(2010)](https://doi.org/10.2140/camcos.2010.5.65).

For theoretical background about Ensemble MCMC (what are the benefits over
simple MCMC? How do they work? What are the pitfalls?), please refer for example to 
[this lecture](https://astrostatistics.psu.edu/su14/lectures/HierarchicalBayesianModelingEnsembleMCMC.pdf)
from Eric B. Ford (Penn State).

## Installation

You can install the stable version of this package from CRAN:

```{r, eval = FALSE}
install.packages("mcmcensemble")
```

or the development version from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("Bisaloo/mcmcensemble")
```

## Usage

```{r}
library(mcmcensemble)

## a log-pdf to sample from
p.log <- function(x) {
    B <- 0.03                              # controls 'bananacity'
    -x[1]^2/200 - 1/2*(x[2]+B*x[1]^2-100*B)^2
}


## use stretch move
res1 <- MCMCEnsemble(p.log, lower.inits=c(a=0, b=0), upper.inits=c(a=1, b=1),
                     max.iter=3000, n.walkers=10, method="stretch")
str(res1)
```

If the [coda](https://cran.r-project.org/package=coda) package is installed,
you can then use the `coda = TRUE` argument to get objects of class `mcmc.list`.
The coda package then allows you to call `summary()` and `plot()` to get 
informative and nicely formatted results and plots:

```{r example-stretch, dev = 'svglite'}
## use stretch move, return samples as 'coda' object
res2 <- MCMCEnsemble(p.log, lower.inits=c(a=0, b=0), upper.inits=c(a=1, b=1),
                     max.iter=3000, n.walkers=10, method="stretch", coda=TRUE)

summary(res2$samples)
plot(res2$samples)
```

```{r example-de, dev = 'svglite'}
## use different evolution move, return samples as 'coda' object
res3 <- MCMCEnsemble(p.log, lower.inits=c(a=0, b=0), upper.inits=c(a=1, b=1),
                     max.iter=3000, n.walkers=10, 
                     method="differential.evolution", coda=TRUE)

summary(res3$samples)
plot(res3$samples)
```

## Parallel processing

This package is set up to allow transparent parallel processing when requested 
by the user thanks to the framework provided by the 
[future](https://cran.r-project.org/package=future) package. To enable parallel 
processing, you must run:

```{r, eval = FALSE}
future::plan("multiprocess")
```

at the start of your session.

## Similar projects
 
The methods used in this package also have (independent) implementations in 
other languages:

- emcee in Python (<https://doi.org/10.21105/joss.01864>)
- gwmcmc in Matlab (<https://github.com/grinsted/gwmcmc>)
